---
title: "Tidymodels Exercise second"
author: "Mateusz Zajda"
format: 
  html:
    toc: true
    toc-title: "contents"
    number-depth: 5
    number-sections: true
    code-fold: show
    code-summary: "Hide/Show code"
    code-tools: true
    code-block-border-left: "black"
    code-line-numbers: true
    code-copy: true
    html-math-method: katex
    embed-resources: true
    smooth-scroll: true
    self-contained: true
    anchor-sections: true
    link-external-icon: true
    link-external-newwindow: true
    theme: 
      light: journal
      dark: darkly
    fontsize:  1.0em
    linestretch: 1.5
date: today
abstract-title: "About"
abstract: "Exercise **second**"
execute: 
  eval: true
  warning: false
  echo: true
  output: true
  error: false
  cache: true

---

```{=html}
<style type="text/css"> body {text-align: justify} </style>
```

## Second exercise
The task of this exercise is to train model for ozone prediction with logistic_regresion.

### Required packages
```{r}
library(tidymodels) 
library(skimr) 
library(GGally) 
library(openair)
library(DT) 
library(dplyr)
tidymodels_prefer()
```

### Data preparation

```{r}

air <- 
  mydata |> 
  selectByDate(year = 2002)
air |> skim()

air <- 
  air |> na.omit()
```

### Changing Ozone to categorical data

```{r}
air <- 
  air |>  
  mutate(ozone = cut(
  o3,
  breaks = c(-0.1, 10, 53),
  labels = c("Niskie", "Wysokie")
  ))

air |> 
  select_if(is.numeric) |> 
  cor(use = "complete.obs") |> as.data.frame() |> select(o3)
```


### Splitting the data

```{r}

set.seed(222)
data_split <- initial_split(data=air, prop = 3/4, strata = ozone)
train_data <- training(data_split)
test_data <- testing(data_split)
```

### Creating recipe for ozone classification model

```{r}
ozone_recipe <- 
  recipe(ozone ~ ., data = train_data) |> 
  update_role(o3, wd, date, pm10, pm25, so2, co, no2, new_role = "ID") |> 
  step_BoxCox(ws, nox, no2) |>  # daje ten sam efet co metoda Yeo-Johnson
  # step_normalize(ws, nox, no2) |> # nie pomaga 
  step_date(date, features = c("month")) |> 
  step_time(date, features = c("hour")) |>
  step_mutate(date_hour = as.factor(date_hour)) |>  
  step_dummy(all_nominal_predictors()) |> 
  step_zv()
```

### Examine the recipe

```{r}
ozone_recipe |> summary()
```
```{r}
ozone_recipe |> 
  prep() |> 
  bake(new_data = train_data) |> 
  head(10) |> 
  DT::datatable()
```

### Defining model

```{r}
lr_mod <-  
  logistic_reg() |> 
  set_engine("glm")
```

### Creating workflow

```{r}
ozone_work <- 
  workflow() |> 
  add_model(lr_mod) |> 
  add_recipe(ozone_recipe)
```

### Fitting data

```{r}
ozone_fit <- 
  ozone_work |> 
  fit(data=train_data)
```

### Examine fitting
```{r}
add_stars <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("")
  }
}
ozone_fit |> 
  extract_fit_parsnip() |> 
  tidy() |> 
  mutate(coef_stars = sapply(p.value, add_stars)) |> 
  print(n = Inf)
```

### Prediction

```{r}
ozone_fit |> 
  predict(test_data, type="prob")
```

### Concat fit with test data
```{r}
pred_test <- 
  augment(ozone_fit, test_data) |>
  select(-ws,
         -wd,
         -o3,
         -nox,
         -no2,
         -pm10,
         -pm25,
         -so2,
         -co,
         -date)
pred_test
```

### Display ROC curve
```{r}
pred_test  |> 
  roc_curve(truth = ozone, .pred_Niskie) |> 
  autoplot()
```
### Counting 'Niskie' and 'Wysokie' values in test set and prediction set
```{r}
pred_test |> count(ozone)
pred_test |> count(.pred_class)
```